---
title: "Pima Indians Diabetes - Spotcheck"
output:
  html_document:
    df_print: paged
---

```{r, include=FALSE}
library(tidyverse)
library(rlang)
library(caret)
library(emc)
library(recipes)
library(forcats)

knitr::opts_chunk$set(fig.width = 14, fig.height = 8, echo = FALSE, warning = FALSE, message = FALSE, comment = "")
options(width = 120)

source("./pid_load.R")
```

# Variable Strategy

## Pregnant

None

## Glucose

Only two missing values. Consider "reasonably complete".
## Pressure

30 missing values. Consider "reasonably complete".

## Mass

9 missing values. Consider "reasonably complete".

## Triceps

185 missing values. Consider "incomplete".

## insulin

301 missing values. Consider "incomplete".

```{r}
quick_spot_df <- pid_load() %>% pid_holdout() %>% pid_clean()
numeric_cols <- quick_spot_df %>% select_if(is.numeric) %>% colnames() %>% syms()
complete_cols <- c("pregnant", "pedigree", "age")
reasonable_complete_cols <- c("glucose", "mass", "pressure")
incomplete_cols <- c("triceps", "insulin")

```

# Quick Spotcheck Algorithms

## Strategy

Standardize variables.

Impute reasonably complete values with median.

Impute incomplete values with bag imputation.

## View algorithms

```{r}

emc_algorithms() %>% filter(two_class, core)

```

Try 

```{r}
quick_spot_algorithms <- c("earth", "glmnet", "lda", "nb", "ada", "rf", "svmLinear", "svmRadial", "xgbLinear", "xgbTree")
paste(quick_spot_algorithms, collapse = ", ") %>% cat()
```

## Spotcheck parameters

View the spotcheck parameters

```{r}
quick_spot_parameters <- 
    expand.grid(algorithm = quick_spot_algorithms) %>% as_tibble() %>%
    mutate(label = sprintf("Quick spotcheck algorithm '%s'", algorithm))
quick_spot_parameters
```

## Spotcheck emc

```{r}
quick_spot_metric <- "Accuracy"
quick_spot_control <- trainControl(method = "repeatedcv", number = 5, repeats = 5)
quick_spot_recipe <- recipe(diabetes ~ ., quick_spot_df) %>% 
    step_meanimpute(reasonable_complete_cols) %>%
    step_scale(all_predictors()) %>%
    step_center(all_predictors()) %>%
    step_BoxCox(all_predictors()) %>%
    step_bagimpute(triceps) %>%
    step_bagimpute(insulin)


if (!exists("prior_quick_spot_emc")) {
    prior_quick_spot_emc <- NULL
}

# prior_quick_spot_emc <- quick_spot_emc

quick_spot_emc <-
    emc_bind(
        quick_spot_parameters,
         emc_record(
             train = train(
                 quick_spot_recipe,
                 data = quick_spot_df,
                 method = algorithm,
                 trControl = quick_spot_control,
                 metric = quick_spot_metric
             ),
             .label = label,
             .verbose = FALSE,
             .cache_path = paste0("Cache/", label, ".rds")
         ),
        emc_performance(train),
        emc_resamples(train)
    )

quick_spot_emc %>% dplyr::select(label, Accuracy, train_error, train_warnings, train_duration) %>% arrange(desc(Accuracy))
```

## Accuracy Boxplot

```{r}

quick_spot_resamples <- 
    quick_spot_emc %>%
    dplyr::select(algorithm, resamples) %>% 
    unnest(resamples)

quick_spot_resamples %>%
    ggplot() +
    geom_boxplot(aes(x = fct_reorder(algorithm, Accuracy), y = Accuracy))
    
```

Ada has a higher median value but I trust earth better because its a linear algorithm and has a more stable accuracy.

## Accuracy Density Plot

```{r}
quick_spot_emc %>% 
    dplyr::select(algorithm, resamples) %>% 
    filter(algorithm != "null") %>%
    unnest(resamples) %>%
    ggplot() +
    geom_density(aes(x = Accuracy, color = algorithm))
```

Possibly non-linear response. SVM might help.

After checking it doesn't! Earth gives a nice stable result. We'll go with that.

We've hit the top accuracy of 77%.

Get the final model.

```{r}
final_model <- quick_spot_emc %>% filter(algorithm == "earth") %>% pull(train) %>% first()
final_model
```

Look like there is room for tuning.

But lets just grab the final model and validate it.

```{r}
final_model <- quick_spot_emc %>% filter(algorithm == "earth") %>% pull(train) %>% first()
test_df <- pid_training(pid_load())
test_predictions <- predict(final_model, newdata = test_df)
caret::confusionMatrix(test_predictions, test_df$diabetes)

```

73% accuracy with the 95% CI inside our 77% target.

# Further Work

' Add features to emc to extract caret resamples.
parallelplot(spotCheck)

' Add features to emc to for pairwise scatter plots
splom(spotCheck)

' Add features to emc to extract caret resamples.
summary(diff(spotCheck))

