---
title: "SciFi Deadly"
author: "Matthew Coad"
date: "4 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)

death_vec <- c(
    1975, "Robert A. Heinlein", 1907, 1988,
    1976, "Jack Williamson", 1908, 2006,
    1977, "Clifford D. Simak", 1904, 1988,
    1978, NA, NA, NA,
    1979, "L. Sprague de Camp",1907 ,2000,
    1980, NA, NA, NA,
    1981, "Fritz Leiber",1910 ,1992,
    1982, NA, NA, NA,
    1983, NA, NA, NA,
    1984, "Andre Norton", 1912, 2005,
    1985, NA, NA, NA,
    1986, "Arthur C. Clarke", 1917, 2008,
    1987, "Isaac Asimov", 1920, 1992,
    1988, "Alfred Bester", 1913, 1987,
    1989, "Ray Bradbury", 1920, 2012,
    1990, NA, NA, NA,
    1991, "Lester del Rey", 1915, 1993,
    1992, NA, NA, NA,
    1993, "Frederik Pohl", 1919, 2013,
    1994, NA, NA, NA,
    1995, "Damon Knight", 1922, 2002,
    1996, "A. E. van Vogt", 1912, 2000,
    1997, "Jack Vance", 1916, 2013,
    1998, "Poul Anderson", 1926, 2001,
    1999, "Hal Clement", 1922, 2003,
    2000, "Brian W. Aldiss", 1925, 2017,
    2001, "Philip Jos ? Farmer", 1918, 2009,
    2002, NA, NA, NA,
    2003, "Ursula K. Le Guin", 1929, NA,
    2004, "Robert Silverberg", 1935, NA,
    2005, "Anne McCaffrey", 1926, NA,
    2006, "Harlan Ellison", 1934, NA,
    2007, "James Gunn", 1923,NA,
    2008, "Michael Moorcock", 1939, NA,
    2009, "Harry Harrison",1925, 2012,
    2010, "Joe Haldeman", 1943, NA,
    2011, NA, NA, NA,
    2012, "Connie Willis", 1945, NA,
    2013, "Gene Wolfe", 1931, NA,
    2014, "Samuel Delany", 1942, NA,
    2015, "Larry Niven", 1938, NA,
    2016, "C.J. Cherryh", 1942, NA,
    2017, "Jane Yolen", 1939, NA
)

death_mtr <- matrix(death_vec, ncol = 4, byrow = TRUE)
gm <- tibble(year = as.integer(death_mtr[, 1]), author = death_mtr[, 2], birth = as.integer(death_mtr[, 3]), death = as.integer(death_mtr[, 4]))

```

## Is being a sci fi author bad for your health?

Quick and dirty data analysis that a perception held by one of my friends that their favorite SciFi authors seem to die young.

Ah. I grabbed the list of hugo award winners from some website, cleaned it up and did some plots.

**Whats the average age hugo award winners died?**

```{r}

gm %>% na.omit() %>% mutate(age_death = death - birth) %>% summarise(avg_d = mean(age_death))

```

**Plot of Hugo award winners death**

```{r}

gm %>% na.omit() %>% mutate(age_death = death - birth) %>% ggplot(aes(x = age_death)) + geom_density()
```

Hugo award winners look like they in general live to a ripe old age.

**Conclusion**: Famous SciFi authors don't die young!
