---
title: "Subspots Exploration"
output:
  html_document:
    self_contained: false    
---

Exploration of the classic sunspots data set.

Challange: Can we detect and quantify the eleven year sunspot cycle.

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
library(tidyverse)
library(rlang)
library(emc)

```

```{r, message=FALSE}

min_year <- 1749
max_year <- 1983
months <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

sunspots_raw <- datasets::sunspots

sunspots_work <- tibble(count = as.numeric(sunspots_raw))
sunspots_work$year <- rep(seq(min_year, max_year), each = 12)
sunspots_work$month <- rep(seq(1, 12), times = max_year - min_year + 1)
sunspots_work$index <- 1:nrow(sunspots_work)

sunspots_work$smooth_count = predict(loess(count ~ index, sunspots_work, span = 0.03))

peak_min <- 1
peak_max <- nrow(sunspots_work)
peak_seq <- seq(peak_min, peak_max)
peak_count <- sunspots_work$smooth_count
peak_interior <- peak_seq > peak_min & peak_seq < peak_max
peak_match <- peak_seq %>% map_lgl(~peak_interior[.] && peak_count[.] > peak_count[. - 1] && peak_count[.] > peak_count[. + 1])
peak_index <- which(peak_match)
peak_smooth_count <- sunspots_work$smooth_count[peak_match]

sunspot_peaks <- tibble(smooth_count = peak_smooth_count, index = peak_index)

cycle_length <- diff(peak_index)
cycle_years <- cycle_length / 12
cycle_start <- peak_index[1:length(peak_index) - 1]
cycle_start_count <- sunspots_work$smooth_count[cycle_start]
cycle_end <- peak_index[2:length(peak_index)]
cycle_end_count <- sunspots_work$smooth_count[cycle_end]
cycle_index <- seq(1, length(peak_index) - 1)

sunspot_cycles <-
    tibble(
           cycle = cycle_index,
           length = cycle_length,
           years = cycle_years,
           start = cycle_start,
           start_count = cycle_start_count,
           end = cycle_end,
           end_count = cycle_end_count)
sunspots_cycle_index <-
    list(sunspot_cycles$start, sunspot_cycles$end, sunspot_cycles$cycle) %>%
    pmap_df(function(start, end, cycle) tibble(cycle = cycle, index = seq(start + 1, end), cycle_index = seq(1, end - start), cycle_length = end - start)) %>%
    mutate(cycle_prop = cycle_index / cycle_length)

sunspots <- sunspots_work %>% left_join(sunspots_cycle_index, by = "index")

sunspots_distribution <- emc_variable_distribution(sunspots)

sunspots_yearly <- sunspots %>% group_by(year) %>% summarise(count = sum(count))

print("Data Loaded")

```

## Variable Statistics

```{r}
sunspots_distribution %>% kable()
```

## Count Exploration

**What does a simple explore of the year vs count show?**
```{r}
ggplot(sunspots, aes(x = year, y = count)) + geom_line()
```

The sunspot cycle is easily visible.

## Yearly Exploration

Summarise the sunspots counted per year. 

**Lets see if the cycle is still visible.**

```{r}
ggplot(sunspots_yearly, aes(x = year, y = count)) + geom_line()
```

The pattern is there.

## Monthly Exploration

**Is the number of sunspots related to the month?**
```{r}
ggplot(sunspots, aes(x = month, y = count)) + geom_point()
```

No readily discernable pattern. It doesn't appear that the sunspot count is related to
the month. 

Which is as expected. However we could expect plenty of cyclic datasets to be connected to the year
so its good to check.

## Cycle Exploration

Lets see if we can't find a way to detect the cycle peaks. Basically we are looking
for a local maxima.

**Zoom in on a few cycles**

```{r}

sunspots %>%
    filter(index <= 500) %>%
    ggplot(aes(x = index, y = count)) +
    geom_line()

```

Its a bit noisy.

**Try loess smoothing**

```{r}

sunspots %>%
    filter(index <= 500) %>%
    ggplot(aes(x = index, y = count)) +
    geom_line(color = "darkgrey") +
    geom_line(aes(y = smooth_count), color = "black")

```

Okay that should make it easy to detect the peaks.

Lets try something simple. 

**Find smoothed point that is bigger than its immediate neighbours. And plot.**

```{r}

sunspots %>%
    ggplot(aes(x = index, y = count)) +
    geom_line(color = "darkgrey")  +
    geom_line(aes(y = smooth_count), color = "black") +
    geom_point(data = sunspot_peaks, aes(x = index, y = smooth_count), color = "Red", size =2)


```

Okay looks like we have detected the peaks.

Number of peaks detected is: `r nrow(sunspot_peaks)`.  
The minimum months between two peaks is: `r min(sunspot_cycles$length)`.  
The maximum months between two peaks is: `r max(sunspot_cycles$length)`.
The mean months between two peaks is: `r mean(sunspot_cycles$length)`.  

Doesn't seem to be any double ups so i think we are good.

## Sunspot Cycle

```{r}

sunspot_cycles %>%
    ggplot(aes(x = years)) +
    geom_density()

```

The mean length of the Sunspot Cycle is: `r mean(sunspot_cycles$years)` years.
The standard deviation of the length of the Sunspot Cycle is: `r sd(sunspot_cycles$years)`.  


## Plot Cycle

Now that we know the sunspot cycles we can plot them over each other.

```{r}

sunspots %>%
    na.omit() %>%
    ggplot(aes(x = cycle_index, y = count, group = cycle, color = cycle)) +
    geom_point(alpha = 0.5, size = .5) + geom_line(aes(y = smooth_count))

```

Or plot them side-by-side.

```{r}

sunspots %>% na.omit(sunspots) %>% ggplot(aes(x = cycle_index, y = count, group = cycle, color = cycle_length)) + geom_point(alpha = 0.5, size = .5) + geom_line(aes(y = smooth_count)) + facet_wrap(~cycle)

```

Interesting. It looks like their is a relationship between the length of a cycle and the strength of the next peak.


## Cycle length vs cycle end height

```{r}

sunspot_cycles %>% ggplot(aes(x = length, y = end_count)) + geom_point() + geom_smooth(method = "lm")

```